---
title: "TO_Shrooms"
author: "Fishbowlz"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
#### Our business: Mushroom Marketplace is an innovative cooperative platform that specializes in the sale of mushrooms, offering a unique marketplace for foragers to connect with a global network of buyers. This platform facilitates a seamless exchange between those who harvest nature’s bounty and those who seek the earthy delights of mushrooms delivered directly from the wilderness to your home.

#### The issue: The increasing trend of mushroom consumption, averaging 3.65 lbs per capita in 2022, is accompanied by the alarming 7,500 poison reports linked to toxic mushrooms, underscoring the critical need for caution and education. This issue is especially prevalent in Michigan, where the majority of these distressing incidents have been reported.

#### Business concerns: To ensure Mushroom Marketplace's success, addressing its challenges is imperative. Providing vetted information on the edibility of mushrooms is crucial, as accurate safety ratings not only reduce liability but also foster trust and encourage repeat patronage from conscientious consumers. This step is key to mitigating concerns and minimizing the chance of adverse reactions, making accurate safety information an integral component in building trust and securing continued support from conscientious consumers.






## Get relevant libraries
```{r}
library(caret)
library(neuralnet)
library(class)
library(kernlab)
library(C50)
library(janitor)
library(tidyr)
library(dplyr)
library(randomForest)
```

## Read Data
```{r}
# Reading the mushroom data from a CSV file and displaying its structure
shrooms <- read.csv("mushrooms.csv", stringsAsFactors = TRUE)
str(shrooms)
```

## Clean Data
```{r}
#in our dataset all veil types are "p" so they were removed
shrooms$veil.type <- NULL

# class is our output variable, so it was made numeric. 1 means poisonous, 0  means edible
shrooms$class <- ifelse(shrooms$class == "p", 1, 0)
#shrooms$stalk.root <- ifelse(shrooms$stalk.root == "?", "missing", shrooms$stalk.root)

colnames(shrooms)[1] ="is_poisonous"

# Load the required libraries
library(caret)
library(zoo)  # for na.aggregate

```

```{r}
# Using model.matrix to convert all the factors to dummy variables
shroomsmm <- as.data.frame(model.matrix(~.-1,shrooms))
str(shroomsmm)
```

```{r}
# 50 - 50 split
set.seed(12345)

shrooms_prop <- 0.5

# Splitting the data into training and testing sets
train_shrooms_rows <- sample(1:nrow(shroomsmm),shrooms_prop*nrow(shroomsmm))
shrooms_train <- shroomsmm[train_shrooms_rows, ]
shrooms_test <- shroomsmm[-train_shrooms_rows, ]
shrooms_train <- na.aggregate(shrooms_train, FUN = mean)
shrooms_test <- na.aggregate(shrooms_test, FUN = mean)

# Creating a function called "normalize" to use the min/max method to normalize the data. This is because a consistent scale is required for ANN and KNN models.
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

shrooms_norm <- as.data.frame(lapply(shroomsmm, normalize))

# Repeating the same exact process with normalized data (which was stored in shrooms_norm) for ANN and KNN.
shrooms_norm_train <- shrooms_norm[train_shrooms_rows, ]
shrooms_norm_test <- shrooms_norm[-train_shrooms_rows, ]

# For KNN, the x and y variables need to be separated from each other. Our y-variable, "is_poisonous," is the 2nd column in the dataset, so we are filtering the columns in our train and test data by that value.
shrooms_norm_train_x <- shrooms_norm_train[, -1]
shrooms_norm_test_x <- shrooms_norm_test[, -1]
shrooms_norm_train_y <- shrooms_norm_train[, 1]
shrooms_norm_test_y <- shrooms_norm_test[, 1]

shrooms_train_no_mm <- shrooms[train_shrooms_rows, ]
shrooms_test_no_mm <- shrooms[-train_shrooms_rows, ]

```

## Logistic Model
#### The confusion matrix for the logistic model reveals that the model achieved perfect accuracy on the test data, correctly classifying all instances as either edible (class 0) or poisonous (class 1). The high sensitivity, specificity, and positive predictive value suggest that the model effectively identified both positive and negative cases, indicating strong performance in distinguishing between edible and poisonous mushrooms in the dataset.
```{r}
logistic_model <- glm(is_poisonous ~ ., data = shrooms_train, family = "binomial")
logistic_pred <- predict(logistic_model, shrooms_test, type = "response")
logistic_pred_binary <- ifelse (logistic_pred >= 0.3, 1, 0)

confusionMatrix(as.factor(logistic_pred_binary), as.factor(shrooms_test$is_poisonous), positive = "1")
```


## Decision Tree Model
##### The confusion matrix for the decision tree model indicates high accuracy, with all instances being correctly identified as either edible (class 0) or poisonous (class 1). The model shows strong sensitivity, specificity, and positive predictive value, reflecting its capability to distinguish between the two classes in the dataset effectively. The performance metrics, along with a Kappa value of 1, imply that the decision tree model has achieved a high level of agreement with the true class labels on the dataset provided.
```{r}
dt_model <- C5.0(as.factor(is_poisonous) ~ ., data = shrooms_train)

plot(dt_model)

dt_pred <- predict(dt_model, shrooms_test)

confusionMatrix(as.factor(dt_pred), as.factor(shrooms_test$is_poisonous), positive = "1")
```


## Random Forest Model
#### The confusion matrix for the random forest model indicates perfect performance, with all instances accurately classified as either edible (class 0) or poisonous (class 1). The model exhibits perfect accuracy, sensitivity, specificity, and positive predictive value, reflecting its exceptional ability to distinguish between mushroom classes in the dataset. The Kappa value of 1 further underscores the optimal agreement between predicted and actual class labels.
```{r}
rf_model <- randomForest(is_poisonous ~ ., data = shrooms_train)
rf_pred <- predict(rf_model,shrooms_test)
rf_pred_binary <- ifelse(rf_pred >= 0.3, 1, 0)
confusionMatrix(as.factor(rf_pred_binary),as.factor(shrooms_test$is_poisonous),positive="1")
```

## KNN Model
#### The KNN model's results were interesting because they were not perfectly classified as with the other models. The reasoning for this is that KNN is sensitive to noisy data and outliers. If there are instances in the dataset that are far from their neighbors, it can lead to misclassifications. The confusion matrix for the KNN model indicates high performance, with only a small number of instances misclassified. The model achieved an accuracy of 99.88%, demonstrating excellent sensitivity, specificity, and positive predictive value. The Kappa value of 0.9975 suggests strong agreement between predicted and actual class labels, highlighting the KNN model's effectiveness in distinguishing between edible and poisonous mushrooms in the dataset.
```{r}
knn_pred <- knn(shrooms_norm_train_x, shrooms_norm_test_x, shrooms_norm_train_y, k =20)
confusionMatrix(as.factor(knn_pred), as.factor(shrooms_norm_test_y), positive = as.character(1))
```


## SVM Vanilladot Model
#### The confusion matrix for the SVM VanillaDot model indicates perfect classification. The model achieved an accuracy of 1, which means it correctly predicted all the instances of both classes (0 and 1) without any errors. The sensitivity and specificity are both 1.0000, signifying that the model has a perfect true positive rate and a perfect true negative rate, respectively.
```{r}

svm_model_vanilla <- ksvm(is_poisonous ~ ., data=shrooms_train , kernel = "vanilladot")
svm_pred_vanilla <- predict(svm_model_vanilla,shrooms_test)
svm_pred_vanilla_binary <- ifelse(svm_pred_vanilla >= 0.3, 1, 0)

# Evaluation
confusionMatrix(as.factor(svm_pred_vanilla_binary),as.factor(shrooms_test$is_poisonous))
```

## ANN Model
#### The confusion matrix for the ANN model shows that it has also achieved flawless performance, similar to all the other models except for the KNN model. The model has an accuracy of 1, indicating that it has correctly classified all predictions for both classes (0 and 1). The sensitivity and specificity values are both at the maximum of 1.0000, meaning the model has perfectly identified all true positives and true negatives without any false positives or false negatives.
```{r}

ann_model <- neuralnet(is_poisonous ~ ., data = shrooms_norm_train, hidden=c(2,2))
ann_pred <- predict(ann_model, shrooms_norm_test)
ann_pred_binary <- ifelse(ann_pred >= 0.3, 1, 0)

# Evaluation
confusionMatrix(as.factor(ann_pred_binary), as.factor(shrooms_norm_test$is_poisonous), positive = as.character(1))

```

## Creating Stacked Model
### Creating Data Frame of Models
```{r}
# because ann and svm predictions output as matrices, we need to convert them into a vector, which we do below
ann_pred_list <- split(ann_pred, row(ann_pred))
ann_pred_vector <- unlist(ann_pred_list)
svm_pred_list <- split(svm_pred_vanilla, row(svm_pred_vanilla))
svm_pred_vector <- unlist(svm_pred_list)

# We are creating a data frame of all the vectors of test data predictions we have done for all 5 different models.
shrooms_preds <- data.frame(log = logistic_pred, 
                          knn = as.numeric(as.character(knn_pred)), 
                          decisiontree = as.numeric(as.character(dt_pred)), 
                          ann = ann_pred_vector, 
                          svm = svm_pred_vector, 
                          true = shrooms_test$is_poisonous)
```

### Breaking Models into Test and Train 

```{r}
# We are now breaking down our predictions into test and train data (another layer). That way, we have something to train our decision tree model of models on. We are using the same process to split the data as performed earlier.
set.seed(12345)
tree_train_rows <- sample(1:nrow(shrooms_preds),.7*nrow(shrooms_preds))
tree_test <- shrooms_preds[-tree_train_rows, ]
tree_train <- shrooms_preds[tree_train_rows, ]
```

### Building the Stacked Model
#### The stacked model combines the predictions of multiple individual models to make a final prediction. In the provided code, predictions from various models (logistic regression, KNN, decision tree, SVM Vanilladot, and ANN) are combined into a new dataset, and a decision tree model is trained on this model. The decision tree in the stacked model learns to weigh and combine the predictions of the base models, aiming to improve overall predictive performance. This approach helps by balancing out any weaknesses in individual models and can lead to a more robust and accurate final prediction.
```{r}
# Build decision tree model, incorporating cost matrix
tree_model <- C5.0(as.factor(true) ~ .,data=tree_train)
```





### Predicting and Evaluating the Model
#### The output of this shows us that only the log_model actually affects the dataset, so we should just use this given model (log_model) in our RShiny application.
```{r}
# Predicting the test data using decision tree model using the "predict" command
tree_predict <- predict(tree_model, tree_test)
# Creating a confusion matrix and plot to evaluate our model of models! I'm storing the confusion matrix in the variable "tree_confusion_matrix" so I can pull out the false positive/false negative values later on.
tree_confusion_matrix <- confusionMatrix(as.factor(tree_predict),as.factor(tree_test$true), positive="1")
print(tree_confusion_matrix)
plot(tree_model)
```
```{r}
str(tree_model)
```

## Building the App's Model
#### Since we are using the Logistical model, the outputs are identical from what is exhibited above. 
```{r}
#Because the apps inputs do not use a model matrixed set of inputs, we re-made the logistic model without the mm function in the train data

log_model_out <- glm(is_poisonous ~ ., data = shrooms_train_no_mm, family = "binomial")

log_out_pred <- predict(log_model_out, shrooms_test_no_mm, type = "response")
log_out_pred_binary <- ifelse (log_out_pred >= 0.3, 1, 0)

confusionMatrix(as.factor(log_out_pred_binary), as.factor(shrooms_test_no_mm$is_poisonous), positive = "1")

saveRDS(log_model_out, "log_model.rds")
```




# Conclusion
#### Our solution: Our solution at Mushroom Marketplace is a testament to our commitment to safety and innovation. By meticulously crafting six distinct models from our training dataset and integrating them into a comprehensive stacked model, we have identified the optimal approach: a logistic regression model that boasts impeccable accuracy (100%) and Kappa (1) statistics. Our dedicated team has developed an rShiny application that allows users to input data, which is then processed by our carefully selected algorithm. This user-facing application serves as the cornerstone of our platform, ensuring that every mushroom processed via our website is verified for safety, thereby keeping our customers safe while emboldening customer trust. 

